### 2023-06-20 paper/sec-dis-pseudodata.tex
sect 2

Table~\ref{tab:integrated_rates} summarises the predicted integrated event yields for the five detectors considered
  in the present analysis, separated into electron neutrinos and antineutrinos
  and muon neutrinos and antineutrinos.
  %
  These event yields are computed from Eq.~(\ref{eq:event_yields}) with the only
  restriction that the momentum transfer and the final-state invariant mass should be restricted
  to the DIS region,
  \be
  \label{eq:DISconditions}
Q^2 \ge 2~{\rm GeV}^2\quad{\rm and} W^2 \ge 12.5~{\rm GeV}^2 \, .
\ee
  %
  As discussed above, for FASER$\nu$ and SND@LHC we assume the integrated luminosity to
  be that of Run III while for the FPF experiments we assume the full HL-LHC
  luminosity (3 ab$^{-1}$).
  %
   The numbers in parenthesis indicate the event rates corresponding to charm
  production for those experiments with heavy flavour tagging capabilities.

@@

By means of tree-level DIS kinematics one can relate the variables describing the neutrino-nucleon
hard-scattering, $(x,Q^2,E_{\nu})$,
with those describing the leptonic and hadronic final state, namely $\lp E_{\ell},\theta_\ell, E_{h} \rp$,
and that correspond to experimentally measured quantities.
%
The mapping between the two sets of variables is 
\bea
E_{\nu} &=& E_\ell + E_h \, ,\nonumber \\
Q^2 &=& 4E_\ell E_{\nu}\sin^2\lp {\theta_\ell/2}\rp  =  4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp \, , \\
x &=& \frac{Q^2}{2m_N(E_{\nu} - E_\ell)} = \frac{4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp}{2m_NE_{h}} \, , \nonumber
\eea
highlighting how the measurement of $E_{\ell}$, $\theta_\ell$, and $E_{h}$ uniquely specify the hard-scattering
kinematics $(x,Q^2,E_{\nu})$ of the event.
%
Therefore, systematic uncertainties in the measurement of these final-state variables will modify
the assignment of each event to a given bin of $x$, $Q^$, and $E_\mu$ introducing a systematic
error in the event yields and hence in the measurement of the double-differential
neutrino scattering cross-section.

We wish to sample over the space of observables, $(E_{l},E_{h},\theta)$, perform  cuts based on detector performance, calculate the event rates, and then smear this sampling according to experimental uncertainties to estimate the uncertainty on the event rate.

Taking FASER${\nu}$2 as an example, we can write the experimental cuts and uncertainties as

\begin{align}
100 < E_l < E_{\nu,{\rm max}} , \delta E_l = 30\% \nonumber \\
100 < E_h < E_{\nu,{\rm max}} , \delta E_h = 50\% \nonumber \\
0 < \theta < \tan^{-1}(0.5) , \delta\theta = 1~{\rm mrad}
\label{fasernu2systematic_errors}
\end{align}

We first generate a MC dataset, $D_0$, over this space with $N = 10^7$ samples and calculate $x,Q^2,E_{\nu}$ for each event, removing samples with $y > 1$. We then integrate according to Eq. \ref{MCintegration} to produce a distribution of events with bins in $x,Q^2,E_{\nu}$. We then smear each sample according to a Gaussian distribution with widths given by Eq.~\ref{fasernu2systematic_errors} to produce a new data set $D_i$ and repeat to produce $M = 10$ event distributions($M>10$ can later be used, but $M = 10$ appears to be stable). For each bin we take the standard deviation to produce the systematic errors per bin (denoted as "N\textunderscore sys\textunderscore errs" in files named "binned\textunderscore sys-events... .txt" \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results}{here}).

For FASER${\nu}$2, the systematic errors typically dominate over statistical errors, at about $10\%$.

@@

>>>>

Table~\ref{tab:integrated_rates} summarises the predicted integrated event yields for the five detectors considered
  in the present analysis, separated into electron neutrinos and antineutrinos
  and muon neutrinos and antineutrinos.
  %
  These event yields are computed from Eq.~(\ref{eq:event_yields}) with the only
  restriction that the momentum transfer and the final-state invariant mass should be restricted
  to the DIS region,
  \be
  \label{eq:DISconditions}
Q^2 \ge 2~{\rm GeV}^2\quad{\rm and}\quad  W^2 \ge 12.5~{\rm GeV}^2 \, .
\ee
  %
  As discussed above, for FASER$\nu$ and SND@LHC we assume the integrated luminosity to
  be that of Run III while for the FPF experiments we assume the full HL-LHC
  luminosity (3 ab$^{-1}$).
  %
   The numbers in parenthesis indicate the event rates corresponding to charm
  production for those experiments with heavy flavour tagging capabilities.

@@

By means of tree-level DIS kinematics, one can relate the variables describing the neutrino-nucleon
hard-scattering, $(x,Q^2,E_{\nu})$,
with those describing the leptonic and hadronic final state, namely $\lp E_{\ell},\theta_\ell, E_{h} \rp$,
and that correspond to experimentally measured quantities.
%
The mapping between the two sets of variables is 
\bea
E_{\nu} &=& E_\ell + E_h \, ,\nonumber \\
Q^2 &=& 4E_\ell E_{\nu}\sin^2\lp {\theta_\ell/2}\rp  =  4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp \, , \\
x &=& \frac{Q^2}{2m_N(E_{\nu} - E_\ell)} = \frac{4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp}{2m_NE_{h}} \, , \nonumber
\eea
highlighting how the measurement of $E_{\ell}$, $\theta_\ell$, and $E_{h}$ uniquely specify the hard-scattering
kinematics $(x,Q^2,E_{\nu})$ of the event.
%
Therefore, systematic shifts in the measurement of these final-state variables will modify
the assignment of each event to a given bin of $x$, $Q^2$, and $E_\mu$, introducing a systematic
error in the event yields and hence in the measurement of the double-differential
neutrino scattering cross-section.

In order to determine these bin-by-bin systematic uncertainties,
we extend the calculation delineated in Eq.~(\ref{sec:pseudo-data_generation}) as follows.
%
Let us consider for illustrative purposes the FASER${\nu}$2 case.
%
From Table~\ref{tab:FPF_experiments}, one reads that the corresponding acceptances
and performance parameters for $E_{\ell}$, $\theta_\ell$, and $E_{h}$   are given by
\bea
E_\ell \ge 100~{\rm GeV} \, , \quad \delta E_\ell \sim 30\% \, ,\nonumber \\
E_h \ge 100~{\rm GeV} \, , \quad \delta E_h \sim 30\% \, ,\\
 \theta_\ell \le \tan^{-1}(0.5) \, , \quad \delta\theta_\ell \sim 1~{\rm mrad} \, . \nonumber 
\label{fasernu2systematic_errors}
\eea
In order to translate the systematic uncertainties $\delta E_\ell$, $\delta E_h $,
and $\delta\theta_\ell$ into systematic errors associated to the binned event yields, 
\be
\label{eq:event_yields_systematic_error}
\delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(i)} \, ,\quad
\delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(i)}
\, ,\quad
\delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(i)} \, ,\qquad i=1,\ldots,N_{\rm bin} \, ,
\ee
we generate first a Monte Carlo set of events, denoted by $\mathcal{D}_0$,
composed by $N_{\rm mc} = 10^7$ samples and determine the assignment of each event
to a point in the $\lp x,Q^2,E_{\nu}\rp$ space.
%
By integrating over this sample by means of Eq.~(\ref{MCintegration}) we determine
the binned event yields for this reference $\mathcal{D}_0$ sample.

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
Provided the number of samples $M$ is large enough, the standard deviation over samples in this $ N_{\rm ev}^{(i)(k)}$
distribution provides the sought-for estimate of the systematic uncertainties.
%
Furthermore, 
since the systematic errors are treated as uncorrelated among them,
by producing samples where only one source of error is varied at a time
we can determine the values of Eq.~(\ref{eq:event_yields_systematic_error}) in each bin.

The end result of the procedure is the estimate of statistical and systematic uncertainties
for each bin of the measurement, from which a experimental covariance matrix can be constructed as
\be
   {\rm cov}_{ij} = \delta_{ij} \lp \delta_{\rm stat}  N_{\rm ev}^{(i)}\rp^2
   + \lp \delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(i)} \rp \lp \delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(j)} \rp 
   + \lp \delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(i)} \rp \lp\delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(j)}\rp 
   + \lp \delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(i)} \rp
   \lp \delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(j)} \rp
   \, ,\qquad
 \nonumber
 \ee
 for $i,j=1,\ldots,N_{\rm bin}$, and the same for the associated correlation
 matrix of the measurement
 \be
 \rho_{ij} =  \frac{{\rm cov}_{ij}}{\sqrt{ {\rm cov}_{ii} }\sqrt{ {\rm cov}_{jj} } } \, . 
 \ee
 The relative covariance matrix, $ {\rm cov}_{ij}/( N_{\rm ev}^{(i)}N_{\rm ev}^{(j)})$, is
 independent of the considered observable and would also apply
 for the double-differential cross-sections Eqns.~(\ref{eq:neutrino_DIS_xsec_FL}) and~(\ref{eq:antineutrino_DIS_xsec_FL}) which are related to the event yields by a constant factor.

 One should emphasize that in a real experiment the actual covariance matrix will be
 composed by a much larger number of uncertainty sources, with typical
  HERA and LHC precision measurements characterised by up to hundreds
 of different sources of systematic error.
 %
 In particular, the assumption that a single source of systematic error, say $\delta E_\ell$,
 is fully correlated among all the bins in $(x,Q^2)$ is likely not to be realistic.
 %
 For this reason, here we present impact results both for the nominal correlation model,
 for the case where systematic and statistical uncertainties are added in quadrature,
 and intermediate scenarios.

 Fig.~\ref{fig:error_plot_FASERv2_14} (central and right panels) display
 the integrated event yields for FASER$\nu$2 as a function of $x$ for only
 systematic errors and for the sum in quadrature of statistical and systematic errors.
 %
 For most of the bins, for the baseline performance assumptions the total systematic
 uncertainty is at the $10\%$ and dominates over the statistical uncertainties.

@@

 In order to generate pseudo-data for double-differential
 neutrino scattering cross-sections at the LHC, we follow the procedure
 used for the HL-LHC PDF projections of~\cite{AbdulKhalek:2018rok} which was
 also adopted in~\cite{Ethier:2021ydt} and~\cite{Greljo:2021kvv} for SMEFT impact projections
 of vector-boson scattering and high-mass Drell-Yan data at the HL-LHC, respectively.
 %
 The starting point are the predictions for the differential neutrino scattering
 cross-section
 \be
 \label{eq:theory_dis_projections}
 \mathcal{O}_i^{{\rm (th)}} \equiv \frac{d^2\sigma^{\nu N}(x_i,Q^2_i,y_i)}{dxdy} \, ,\quad
 i=1,\ldots,N_{\rm bin} \, ,
 \ee
 with $(x_i,Q^2_i,y_i)$ being the corresponding bin centers.
 %
 The observables $\mathcal{O}_i $ are evaluated using {\sc\small YADISM} and {\sc\small PineAPPL}
 as described in Sect.~\ref{sec:settings}.
 %
 An important point here is the choice of PDF set entering the evaluation of
 the observable in Eq.~(\ref{eq:theory_dis_projections}): it should be consistent
 with that of the fitting framework used to assess their impact on the (n)PDFs.
 %
 For instance, when using the {\sc\small xFitter} profiling of PDF4LHC21, one needs
 to generate LHC neutrino pseudo-data also using PDF4LHC21 as input.
 %
 In order words, the generated pseudo-data should be fully consistent with the prior PDF
 set used as baseline, else one is introducing artificially inconsistencies which
 compromise the validity of the projection studies.

 The central values for the LHC neutrino scattering pseudo-data, denoted
 by $\mathcal{O}_i^{{\rm exp}} $, are obtained
 by fluctuating this reference theory prediction by the statistical and systematic
 uncertainties, that is
 \begin{equation}
  \label{eq:pseudo_data}
  \mathcal{O}_i^{{\rm (exp)}}  =   \mathcal{O}_i^{{\rm (th)}}\left( 1+ r^{(i)}_{\rm stat}\delta_{i,{\rm stat}}
  + \sum_{k=E_\ell, E_h,\theta_\ell} r_{{\rm sys},k} \,\delta_{i,{\rm sys}}^{(k)}\right) \, , \qquad i=1,\ldots,N_{\rm bin} \, ,
 \end{equation}
 with $r^{(i)}_{\rm stat}$ and $r_{{\rm sys},k}$ being univariate Gaussian random numbers,
 and $\delta_{\rm stat}$ ($\delta_{\rm sys}^{(k)}$) indicate the relative statistical (systematic)
 uncertainties associated to the $i$-th bin.
 %
 Note that a given systematic variation is assumed to be fully correlated among all the bins
 in the measurement.

 Alternatively, we can add all uncertainties in quadrature and account for a possible
 improvement in the systematic errors with respect to the current estimates,
 \be
 \delta_{{\rm tot},i}^{\rm exp} = \left( \left( \delta_i^{\rm stat}\right)^2 + \sum_{k=1}^{n_{\rm sys}}
\left( f_{{\rm red},j}\delta_{i,j}^{\rm sys} \right)^2\right)^{1/2} \, ,
 \ee
 with $f_{{\rm red},j} \le 1$ quantifies the possible reduction of systematic uncertainties, and
 then the pseudo-data is evaluated by means of
 \begin{equation}
  \label{eq:pseudo_data_v2}
  \mathcal{O}_i^{{\rm (exp)}}  =   \mathcal{O}_i^{{\rm (th)}}\left( 1+ r^{(i)}_{\rm stat}\delta_{i,{\rm tot}}
  \right) \, , \qquad i=1,\ldots,N_{\rm bin} \, .
 \end{equation}
 While the actual correlation model will lie somewhere between no correlation and full correlation,
 these two approaches to generate the pseudo-data bracket the projected impact
 on the PDFs and its dependence on the experimental correlation model.
